---
title: "Investigating COVID-19 Virus Trends"
author: "Eriz Tolay"
date: "9/10/2021"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
library(tidyverse)
```

## COVID-19

A pneumonia of unknown cause detected in Wuhan, China was first internationally reported from China on 31 December 2019. Today we know this virus as Coronavirus. COVID-19 which stands for COronaVIrus Disease is the disease caused by this virus. Since then, the world has been engaged in the fight against this pandemic. Several measures have therefore been taken to "flatten the curve". We have consequently experienced social distancing and many people have passed away as well.

Our analysis tries to provide an answer to this question: **Which countries have had the highest number of positive cases against the number of tests?**

```{r}
# Read the data
covid_df <- read_csv('covid19.csv')
```

```{r}
# Determining the dimensions
dim(covid_df)
```

```{r}
# Assign column names to a variable and display
vector_cols <- colnames(covid_df)
vector_cols
```

```{r}
# Display fist 5 rows
head(covid_df)
```
```{r}
# Display summary of the dataset
glimpse(covid_df)
```

## Filtering and Extracting the Data

Looking at the few lines of our dataset we displayed in the previous step, we can see that the Province_State column mixes data from different levels: country level and state/province level. Since we cannot run an analysis on all these levels at the same time, we need to filter what we are interested in.

We will, therefore, extract only the country-level data in order to not bias our analyses.

```{r}
# To filter for All States in Province_State column
covid_df_all_states <- covid_df %>%
  filter(
    Province_State == "All States"
  )
# Dropping Province_State column
covid_df <- covid_df[,-5]
```

Our analysis would be biased if we made the mistake of comparing a column containing cumulative data and another one containing only one-day data. 

Thereafter, we work mainly with daily data. So we will extract the columns related to the daily measures.

```{r}
covid_df_all_states_daily <- covid_df_all_states %>%
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
```

We will summarize the covid_df_all_states_daily dataframe by computing the overall number of tested, positive, active and hospitalized cases. Then, we can arrange this aggregated data by the overall number of tested cases. Finally, we can extract the first ten rows as the top ten tested cases countries.

```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>%
  group_by(Country_Region) %>%
  summarise(
    tested = sum(daily_tested),
    positive = sum(daily_positive),
    active = sum(active),
    hospitalized = sum(hospitalizedCurr)
  ) %>%
  arrange(-tested)
  
covid_df_all_states_daily_sum
```

```{r}
covid_top_10 <- head(covid_df_all_states_daily_sum,10)
covid_top_10
```

Our goal now is to answer this question: **Which countries have had the highest number of positive cases against the number of tests?**

```{r}
countries <- covid_top_10$Country_Region
tested_cases<- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized

names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries

positive_cases / tested_cases
positive_tested_top_3 <- c("United Kingdom" = 0.11, "United States" = 0.10, "Turkey" = 0.08)
positive_tested_top_3
```

Our goal now is to find a way to keep all the information available for the top three countries that have had the highest number of positive cases against the number of tests.

To make sure we won't lose other information about these countries we can create a matrix that contains the ratio and the overall number of COVID-19 tested, positive, active and hospitalized cases.

```{r}
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

covid_mat <- rbind(united_kingdom, united_states, turkey)
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
covid_mat
```

```{r}
question <- "Which countries have had the highest number of positive cases against the number of tests?"

answer <- c("Positive tested cases" = positive_tested_top_3)
```

```{r}
datasets <- list(
  original = covid_df,
  allstates = covid_df_all_states,
  daily = covid_df_all_states_daily,
  top_10 = covid_top_10
)
matrices <- list(covid_mat)
vectors <- list(vector_cols, countries)
data_structure_list <- list("dataframe" = datasets, "matrix" = matrices, "vector" = vectors)
covid_analysis_list <- list(question, answer, data_structure_list)
covid_analysis_list[[2]]
```
